Bioinformatics analysis of bisulfite-sequencing data

### 2014-11-08 Rmarkdown/biology_background.Rmd
Fixed latex errors and missing references.

The readers of DNA methylation recognise methylated DNA. These readers can recruit additional proteins to the site of the methylated cytosine to perform a variety of functions related to gene expression. For example, the methyl-CpG-binding domain (_MBD_) group of proteins bind to DNA containing a methylated CpG, which can then suppress gene expression by preventing transcription factor binding at that site \cite{Nan:1993gm}. Another group, the ubiquitin-like, containing PHD and RING finger domain (_UHRF_) proteins, help _DNMT1_ methylate hemimethylated DNA, such as the daughter strand created during DNA replication \cite{Sharif:2007ft, Bostick:2007jh}.

@@

@@

oxBS-seq specifically measures 5mC. The input DNA is oxidated by potassium perruthenate (RKRuO_{4}$), which converts 5hmC to 5fC, prior to bisulfite-treatment. Only 5mC is protected from conversion during the bisulfite-treatment, which effectively means that only 5mC remains to be detected at the sequencing stage. The level of 5hmC can be estimated by performing traditional bisulfite-sequencing and then "subtracting" the oxBS-seq signal (5mC) from the bisulfite-sequencing signal (5mC + 5hmC).

@@

## General TODOs
* Nucleotide, residue, (nucleo)base; what's the difference and which term should I use?
	* A nucleotide is composed of a nucleobase (also termed a nitrogenous base), a five-carbon sugar (either ribose or 2-deoxyribose), and one or more phosphate groups \[Coghill, Anne M.; Garson, Lorrin R., ed. (2006). The ACS style guide: effective communication of scientific information (3rd ed.). Washington, D.C.: American Chemical Society. p. 244. ISBN 978-0-8412-3999-9.)\]
* Should I give more detail on Rafa's CGI HMM?
* Define eukaryotes, prokaryotes and viruses
* Hyphenation or not of "bisulfite-sequencing"?
  - Yes, hyphenate.
* General term for methylC-seq, BS-seq, etc.
  - "whole-genome bisulfite-sequencing (WGBS)"
* Is it most accurate to say that restriction enzymes "cleave", "cut" or "digest" DNA?
* Define/describe DNA sequencing and the genomics revolution in the section describing DNA (or reference genome section?)
	* Define "library"
* Fix up reference formating
* Note that when I refer to a read I mean either a single-end read or a paired-end read. I will use to read_1 and read_2 if I want to refer a particular read that makes up a paired-end read.
* \cite{Zhang:2013uu} credit Laurent et al. (2010) with BS-seq and Cokus with WGBS.

>>>>

The readers of DNA methylation recognise methylated DNA. These readers can recruit additional proteins to the site of the methylated cytosine to perform a variety of functions related to gene expression. For example, the methyl-CpG-binding domain (_MBD_) group of proteins bind to DNA containing a methylated CpG, which can then suppress gene expression by preventing transcription factor binding at that site \cite{Nan:1993gma}. Another group, the ubiquitin-like, containing PHD and RING finger domain (_UHRF_) proteins, help _DNMT1_ methylate hemimethylated DNA, such as the daughter strand created during DNA replication \cite{Sharif:2007ft, Bostick:2007jh}.

@@

__TODO: Is it most accurate to say that restriction enzymes "cleave", "cut" or "digest" DNA?__

@@

oxBS-seq specifically measures 5mC. The input DNA is oxidated by potassium perruthenate ($RKRuO_{4}$), which converts 5hmC to 5fC, prior to bisulfite-treatment. Only 5mC is protected from conversion during the bisulfite-treatment, which effectively means that only 5mC remains to be detected at the sequencing stage. The level of 5hmC can be estimated by performing traditional bisulfite-sequencing and then "subtracting" the oxBS-seq signal (5mC) from the bisulfite-sequencing signal (5mC + 5hmC).

@@

### 2014-11-08 Rmarkdown/wgbs_analysis.Rmd
Fixed latex errors and missing references.

---
title: "Analysis of bisulfite-sequencing data"
author: "Peter Hickey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: ../latex/header.tex
  html_document:
    keep_md: true
bibliography: ../latex/phd_thesis.bib
---

>>>>

---
title: "Bioinformatics analysis of bisulfite-sequencing data"
author: "Peter Hickey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: ../latex/header.tex
  html_document:
    keep_md: true
bibliography: ../latex/phd_thesis.bib
---
